The value of intermediation

### 2023-04-09 src/paper/theory.tex
Improve multiple platforms section

\subsection{Multiple platforms}

Imagine that, instead of just one, there are $m$ players of type $P$, and they are perfect substitutes of each other.
That is, the fringe players still need a platform to achieve any value, but it does not matter which or how many platforms are available for them.
Formally, the value function for coalition $S$ is the following:
\begin{align*}
    v(S) = \begin{cases}
        0                              & \text{if } n_P(S) = 0 \\
        f\left(\frac{n_A(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}

Let us start by establishing monotonicity for this version of the model, too.
\begin{proposition}
    The game is monotone if $f$ is increasing and $f(0) \geq 0$.
\end{proposition}
% \textcolor{red}{Note: I have to think more about how to describe this situation in the language of cooperative game theory. The issue is that, for example, the value of $\{P_1, A_1\}$ depends on whether $\{P_2, A_2\}$ also decides to form a coalition and enter the market.}

As before, the main proposition in this section provides us with an expression of the limit of the value for this variant of the game.
\begin{proposition}
    \label{prop:multiple_platforms}
    The limit of the Shapley-value (as $n \to \infty$) of each player of type $P$ is
    \begin{align*}
        \varphi_{P_i}^{\infty, m} = \int_0^1 (1-t) ^ {m-1} f(t) \dt .
    \end{align*}
\end{proposition}

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1.1, samples at={0, 0.02, 0.04, ..., 0.98, 1},
                xtick={0, 1}, ytick={0, 1}, axis lines=left, xlabel={$t$}, legend pos=north west]
                \addplot[ultra thick] {1};
                \addplot[ultra thick, dashed] {1-x};
                \addplot[ultra thick, dotted] {(1-x)^2};
            \end{axis}
        \end{tikzpicture}
        \caption{$(1 - t) ^ {m - 1}$}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1.1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick={0, 1}, axis lines=left, xlabel={$t$}, legend pos=north west]
                \addplot[ultra thick] {x};
                \addplot[ultra thick, dashed] {x - x^2/2};
                \addplot[ultra thick, dotted] {x - x^2 + x^3/3};
                \legend{$m = 1$, $m = 2$, $m = 3$}
            \end{axis}
        \end{tikzpicture}
        \caption{$\int_0^t (1 - s) ^ {m - 1} \ds$}
    \end{subfigure}
    \caption{Illustration of the weighting function for various values of $m$ in the multiple platform case.}
    \label{fig:multiple_platforms}
\end{figure}

@@

The following corollary again demonstrates that the predicted outcomes confirm intuitions.
As the major players become more and more substitutable, the portion of the surplus they can capture decreases, and the aggregated value of the fringe increases.

@@

\subsection{Multiple platforms}

@@

It can be shown that the values of each type of player are rather similar to the one-platform case.
\begin{proposition}
    The unit Shapley-values of the various types of players are
    \begin{align*}
        \varphi_{P_i}(k, \infty) & = \int_0^1 (1-t)^{k-1} f(t, \dots, t) \dt                                 \\
        \varphi_{A_l}(k, \infty)     & = \int_0^1 [1 - (1-t)^k] \partial_l f(t, \dots, t) \dt.
    \end{align*}
\end{proposition}

Similarly, as the next corollary demonstrates, the combined value of platforms is still decreasing in their number.
\begin{corollary}
    Let $\varphi_{P}(k, \infty) = k\varphi_{P_i}(k, \infty)$ denote the aggregated Shapley-values of players of type $P$.
    Then $\varphi_{P}(k, \infty)$ is decreasing in $k$ and $\varphi_{A_l}(k, \infty)$ are increasing in $k$ unless $f$ is constant.
\end{corollary}

>>>>

\subsection{Multiple big players}

Now imagine that, instead of just one, there are $m$ players of type $P$, and they are perfect substitutes of each other.
That is, the fringe players still need one at least of them to achieve any value, but it does not matter which and how many of them are present.
Formally, the value function for coalition $S$ becomes the following:
\begin{align*}
    v(S) = \begin{cases}
        0                              & \text{if } n_P(S) = 0 \\
        f\left(\frac{n_A(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}

Let us start by establishing monotonicity for this version of the model, too.
\begin{proposition}
    The game is monotone if $f$ is increasing and $f(0) \geq 0$.
\end{proposition}
On the other hand, superadditivity is not as simple in this case, as before.
For example, when $f$ is concave, then $m$ coalitions with one platform each and the fringe divided equally among them clearly achieve a higher total payoff than the grand coalition.
Therefore, the results of this section are most applicable to settings when multiple coalitions cannot be formed at the same time for some reason.
For example, \textcite{hart1996bargaining} propose an interpretation when there is a single, indivisible, non-replicable resource or technology (not captured in the value function) necessary to produce any value.

As before, the main proposition in this section provides us with an expression of the limit of the value for this variant of the game.
Notably, as in the case of the weighted value, the share of player $P$ can again be expressed as an integral of $f$ and some weighting\footnote{
    Weighting is not a completely precise term, as $h(t)$ does not integrate to one, and is therefore not a probability distribution in this case.
    Nevertheless, it highlights the parallel between multiple big players and weighted values.
} function $h(t)$, with the latter depending only on the number of big players.
\begin{proposition}
    \label{prop:multiple_platforms}
    The limit of the Shapley-value (as $n \to \infty$) of each player of type $P$ is
    \begin{align*}
        \varphi_{P_i}^{\infty, m} = \int_0^1 h(t) f(t) \dt,
    \end{align*}
    where $h(t) = (1-t) ^ {m-1}$.
\end{proposition}

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1.1, samples at={0, 0.02, 0.04, ..., 0.98, 1},
                xtick={0, 1}, ytick={0, 1}, axis lines=left, xlabel={$t$}, legend pos=north west]
                \addplot[ultra thick] {1};
                \addplot[ultra thick, dashed] {1-x};
                \addplot[ultra thick, dotted] {(1-x)^2};
            \end{axis}
        \end{tikzpicture}
        \caption{$h(t)$}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.8]
            \begin{axis}[xmin=0, xmax=1, ymin=0, ymax=1.1, samples at={0, 0.02, ..., 0.98, 1},
                xtick={0, 1}, ytick={0, 1}, axis lines=left, xlabel={$t$}, legend pos=north west]
                \addplot[ultra thick] {x};
                \addplot[ultra thick, dashed] {x - x^2/2};
                \addplot[ultra thick, dotted] {x - x^2 + x^3/3};
                \legend{$m = 1$, $m = 2$, $m = 3$}
            \end{axis}
        \end{tikzpicture}
        \caption{$\int_0^t h(s) \ds$}
    \end{subfigure}
    \caption{Illustration of the weighting function for various values of $m$ in the multiple big player case.}
    \label{fig:multiple_platforms}
\end{figure}

@@

The next corollary again demonstrates that the predicted outcomes confirm intuition.
As the major players become more and more substitutable, the portion of the surplus they can capture decreases, and the aggregated value of the fringe increases (\cref{fig:multiple_platforms}).

@@

\subsection{Multiple big players}

@@

It can be shown that the values of each type of player are rather similar to the single big player case.
\begin{proposition}
    The unit Shapley-values of the various types of players are
    \begin{align*}
        \varphi_{P_i}(k, \infty) & = \int_0^1 (1-t)^{k-1} f(t, \dots, t) \dt                                 \\
        \varphi_{A_l}(k, \infty)     & = \int_0^1 [1 - (1-t)^k] \partial_l f(t, \dots, t) \dt.
    \end{align*}
\end{proposition}

Similarly, as the next corollary demonstrates, the combined value of the big palyers is still decreasing in their number.
\begin{corollary}
    Let $\varphi_{P}(k, \infty) = k\varphi_{P_i}(k, \infty)$ denote the aggregated Shapley-values of players of type $P$.
    Then $\varphi_{P}(k, \infty)$ is decreasing in $k$ and $\varphi_{A_l}(k, \infty)$ are increasing in $k$ unless $f$ is constant.
\end{corollary}
